\textbf{Microbiota predict \emph{Clostridium difficile} severity in germ-free mice colonized with human feces}

### 2022-04-25 submission/manuscript.tex
incorporate PDS feedback in revision

\textbf{\emph{C. difficile} is able to infect germ-free mice colonized
with human fecal microbial communities without antibiotics.} To produce
gut microbiomes with greater variation than those found in conventional
mouse colonies, we colonized germ-free mice with bacteria from human
feces (17). We inoculated germ-free C57BL/6 mice with homogenized feces
from each of 15 human fecal samples via oral gavage. These human fecal
samples were selected because they represented diverse community
structures based on community clustering (18). The gut communities were
allowed to colonize for two weeks post-inoculation (19). We then
surveyed the bacterial members of the gut communities by 16S rRNA gene
sequencing of murine fecal pellets (Figure 1A). The bacterial
communities from each mouse grouped more closely to those communities
from mice that received the same human fecal donor community than to the
mice who received a different human fecal donor community (Figure 1B).
The communities were primarily composed of populations of
\emph{Clostridia}, \emph{Bacteroidia}, \emph{Erysipelotrichia},
\emph{Bacilli}, and \emph{Gammaproteobacteria}. However, the gut
bacterial communities of each donor group of mice harbored unique
relative abundance distributions of the shared bacterial classes.

@@

\textbf{Statistical analysis and modeling.} To compare community
structures, we calculated Yue and Clayton dissimilarity matrices
(\(\theta\)\textsubscript{YC}) in mothur (84). For this calculation, we
averaged of 1000 sub-samples, or rarified, samples to 2,107 sequence
reads per sample to limit uneven sampling biases. We tested for
differences in individual taxonomic groups that would explain the
outcome differences with LEfSe (85) in mothur (default parameters, LDA
\textgreater{} 4). We tested for differences in temporal trends through
fitting a linear model to each OTU and testing for differences between
histopathological summary scores with LEfSe (85) in mothur (default
parameters, LDA \textgreater{} 3). Remaining statistical analysis and
data visualization was performed in R (v4.0.5) with the tidyverse
package (v1.3.1). We tested for significant differences in
\(\beta\)-diversity (\(\theta\)\textsubscript{YC}), histopathological
scores, and toxin activity using the Wilcoxon rank sum test,
non-unimodality to non-moribund histopathological summary score using
Hartigans' dip test, and toxin detection in mice using the Pearson's
Chi-square test. We used Spearman's correlation to identify which OTUs
that had a correlation between their relative abundance and the
histopathologic summary score. \emph{P} values were then corrected for
multiple comparisons with a Benjamini and Hochberg adjustment for a type
I error rate of 0.05 (86). We built L2 logistic regression models using
the mikropml package (87). Sequence counts were summed by taxonomic
ranks from day 0 samples, normalized by centering to the feature mean
and scaling by the standard deviation, and features positively or
negatively correlated were collapsed into a single feature. We ran our
models using alpha = 0 and lambda values of 1e-0, 1e1, 1e2, 2e2, 3e2,
4e2, 5e2, 6e2, 7e2, 8e2, 9e2, 1e3, 1e4 with a split of 80\% of the data
for training and 20\% of the data for testing. Lastly, we did not
compare murine communities to donor community or clinical data because
germ-free mice colonized with non-murine fecal communities have been
shown to more closely resemble the murine communities than the donor
species community (88). Furthermore, it is not our intention to make any
inferences regarding human associated bacteria and their relationship
with human CDI outcome.

@@

Conceptualization: N.A.L., A.M.S., K.J.F., P.D.S.; Data curation:
N.A.L., K.J.F.; Formal analysis: N.A.L., K.J.F., J.L.L., I.L.B.;
Investigation: N.A.L., A.M.S., H.S., I.L.B., V.B.Y., P.D.S.;
Methodology: N.A.L., A.M.S., K.J.F., J.L.L., H.S., I.L.B., V.B.Y.,
P.D.S.; Resources: N.A.L., A.M.S., P.D.S.; Software: NAL; Visualization:
N.A.L., K.J.F., P.D.S.; Writing - original draft: N.A.L.; Writing -
review \& editing: N.A.L., A.M.S., K.J.F., J.L.L., H.S., I.L.B., V.B.Y.,
P.D.S.; Funding acquisition: V.B.Y.; Project administration: P.D.S.;
Supervision: P.D.S.

>>>>

\textbf{\emph{C. difficile} is able to infect germ-free mice colonized
with human fecal microbial communities without antibiotics.} To produce
gut microbiomes with greater variation than those found in conventional
mouse colonies, we colonized germ-free mice with bacteria from human
feces (17). We inoculated germ-free C57BL/6 mice with homogenized feces
from each of 15 human fecal samples via oral gavage. These human fecal
samples were selected because they represented diverse community
structures based on community clustering (18). After the gut communities
had colonized for two weeks, we confirmed them to be \emph{C. difficile}
negative by culture (19). We then surveyed the bacterial members of the
gut communities by 16S rRNA gene sequencing of murine fecal pellets
(Figure 1A). The bacterial communities from each mouse grouped more
closely to those communities from mice that received the same human
fecal donor community than to the mice who received a different human
fecal donor community (Figure 1B). The communities were primarily
composed of populations of \emph{Clostridia}, \emph{Bacteroidia},
\emph{Erysipelotrichia}, \emph{Bacilli}, and \emph{Gammaproteobacteria}.
However, the gut bacterial communities of each donor group of mice
harbored unique relative abundance distributions of the shared bacterial
classes.

@@

\textbf{Statistical analysis and modeling.} To compare community
structures, we calculated Yue and Clayton dissimilarity matrices
(\(\theta\)\textsubscript{YC}) in mothur (84). For this calculation, we
averaged of 1000 sub-samples, or rarified, samples to 2,107 sequence
reads per sample to limit uneven sampling biases. We tested for
differences in individual taxonomic groups that would explain the
outcome differences with LEfSe (85) in mothur (default parameters, LDA
\textgreater{} 4). We tested for differences in temporal trends through
fitting a linear model to each OTU and testing for differences between
histopathological summary scores with LEfSe (85) in mothur (default
parameters, LDA \textgreater{} 3). Remaining statistical analysis and
data visualization was performed in R (v4.0.5) with the tidyverse
package (v1.3.1). We tested for significant differences in
\(\beta\)-diversity (\(\theta\)\textsubscript{YC}), histopathological
scores, and toxin activity using the Wilcoxon rank sum test,
non-unimodality to non-moribund histopathological summary score using
Hartigans' dip test, and toxin detection in mice using the Pearson's
Chi-square test. We used Spearman's correlation to identify which OTUs
that had a correlation between their relative abundance and the
histopathologic summary score. \emph{P} values were then corrected for
multiple comparisons with a Benjamini and Hochberg adjustment for a type
I error rate of 0.05 (86). We built L2 logistic regression models using
the mikropml package (87). Sequence counts were summed by taxonomic
ranks from day 0 samples, normalized by centering to the feature mean
and scaling by the standard deviation, and features positively or
negatively correlated were collapsed into a single feature. For each L2
logistic regression model, we ran 100 random iterations using values of
1e-0, 1e1, 1e2, 2e2, 3e2, 4e2, 5e2, 6e2, 7e2, 8e2, 9e2, 1e3, 1e4 for the
L2 regularization penalty with a split of 80\% of the data for training
and 20\% of the data for testing. Lastly, we did not compare murine
communities to donor community or clinical data because germ-free mice
colonized with non-murine fecal communities have been shown to more
closely resemble the murine communities than the donor species community
(88). Furthermore, it is not our intention to make any inferences
regarding human associated bacteria and their relationship with human
CDI outcome.

@@

Conceptualization: N.A.L., A.M.S., K.J.F., P.D.S.; Data curation:
N.A.L., K.J.F.; Formal analysis: N.A.L., K.J.F., J.L.L., I.L.B.;
Investigation: N.A.L., A.M.S., H.S., I.L.B., V.B.Y., P.D.S.;
Methodology: N.A.L., A.M.S., K.J.F., J.L.L., H.S., I.L.B., V.B.Y.,
P.D.S.; Resources: N.A.L., A.M.S., P.D.S.; Software: NAL; Visualization:
N.A.L., K.J.F., P.D.S.; Writing - original draft: N.A.L.; Writing -
review \& editing: N.A.L., A.M.S., K.J.F., J.L.L., H.S., I.L.B., V.B.Y.,
P.D.S.; Funding acquisition: V.B.Y.; Project administration: P.D.S.;
Supervision: P.D.S. \newpage
